Longitudinal Manifold Estimation

### 2023-04-19 paper/main.tex
paper updates

@@

\textit{Initialization}: We begin by considering $D$-dimensional observations $\left\{y_{it}\right\}_{i=1, t=1}^{I_t, T}$. Like PME, the LPME algorithm begins with a data reduction step, using the HDMDE method detailed in \cite{mengPrincipalManifoldEstimation2021} to estimate a reduced set of cluster centers from $\left\{y_{it}\right\}_{i=1, t=1}^{I_t, T}$. HDMDE is run separately for each individual time point to ensure the time dimension is preserved, while the number of clusters being estimated is allowed to vary for each time point to accommodate potential changes in the shape of the underlying structure. The centers estimated in this step are then used to reach an initial low-dimensional parameterization.

@@

\textit{Fitting}:

\textit{Tuning}:

>>>>

\section*{Abstract}

@@

\subsection*{Initialization}
We begin by considering $D$-dimensional observations $\left\{y_{it}\right\}_{i=1, t=1}^{I_t, T}$. Like PME, the LPME algorithm begins with a data reduction step, using the HDMDE method detailed in \cite{mengPrincipalManifoldEstimation2021} to estimate a reduced set of cluster centers from $\left\{y_{it}\right\}_{i=1, t=1}^{I_t, T}$. HDMDE is run separately for each individual time point to ensure the time dimension is preserved, while the number of clusters being estimated is allowed to vary for each time point to accommodate potential changes in the shape of the underlying structure. The centers estimated in this step are then used to reach an initial low-dimensional parameterization.

@@

\subsection*{Fitting}
Once the initial parameterization for each cluster center is obtained, the model fitting process begins by using this parameterization as the initialization for runs of the PME algorithm at each time point. Specifically, for each time point, the cluster centers associated with the given time value, their weights, and the corresponding parameterizations for these points are used to initialize a PME run. This results in an estimated principal manifold for each time point. The remainder of the fitting stage involves estimating a thin plate spline function mapping from the time dimension to the coefficient space that smooths over the coefficients for these principal manifolds, given a tuning parameter $\gamma$.

At each time point, the estimated principal manifold has the functional form of a smoothing spline or thin plate spline, depending on the intrinsic dimension $d$. As detailed by \cite{greenSilverman1994}, smoothing splines, and thin plate splines as their higher dimensional analogues, may be considered to have knots at the points being used to fit them. This means that, in order for the coefficients of the manifolds estimated at each time point to be comparable, the parameter values used as inputs when the functions were fit must be identical. Hence, to obtain coefficients that can be compared across time points, a grid of values ranging from the minimum to the maximum of the parameters in each of the $d$ dimensions is input into the estimated embedding functions for each time point $f_t^*$. This yields a set of outcomes corresponding to same parameter values at each time point. Following the discussion in Chapter 7 of \cite{greenSilverman1994}, comparable spline coefficients $\delta$ and $\mathbf{a}$ can then be calculated by minimizing
\[%
  S(g) = (\mathbf{X} - E\delta - T^{T}\mathbf{a})^{T}(\mathbf{Y} - E\delta - T^{T}\mathbf{a}) + \gamma\delta^{T}E\delta
,\]%
where matrix $E$ is defined by $E_{ij} = \eta_{d}(\|\mathbf{r}_i - \mathbf{r}_j\|)$ and matrix $T$ has $T_{ij} = \phi_i(\mathbf{r}_j)$. This minimization is conducted by solving
\[%
  \left[
    \begin{array}{cc}
      E + \gamma I & T^{T} \\
      T^{T} & 0
    \end{array}
  \right]\left(
    \begin{array}{c}
      \mathbf{\delta} \\
      \mathbf{a}
    \end{array}
  \right) = \left(
    \begin{array}{c}
      \mathbf{Y} \\
      0
    \end{array}
  \right)
.\]%

\subsection*{Tuning}
